---
title: "Analysis workflow"
author: "Pep Porrà"
date: "May 9, 2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Goal

To investigate tools to implement a DS analysis workflow that makes as easy as possible to start an analysis in `R` that:

1. It is sharable
1. It is reproducible
1. It is extensible

## Tools and Packages

### `installr` package

It makes it easier to upgrade `R` and other packages, for instance RTools [link](https://cran.r-project.org/web/packages/installr/index.html)

```{r}
library(installr)
```

### `RTools`

`RTools` package is required to build packages that work in Windows [link](https://cran.r-project.org/bin/windows/Rtools/history.html). A new version exist for `R` 4.0.0 [link](https://cran.r-project.org/bin/windows/Rtools/)

We first check if `RTools` is installed using the `find_rtools` function in `devtools` package


```{r}
library(devtools)
```

```{r}
find_rtools()
```
```{r}
warning_1 <- 'WARNING: Rtools is required to build R packages, but is not currently installed.
Please download and install Rtools custom from http://cran.r-project.org/bin/windows/Rtools/.'
```

fucntion `install_Rtools(check = TRUE)` does not work. I downloaded version 3.5 directly from [RTools](https://cran.r-project.org/bin/windows/Rtools/history.html). Download the version required and execute it. I have chosen to add it to path and not to add it to registry.


### R version info

We can obatin info about the session we are in with the function in the package `utils`

```{r}
R.version
```

### R session info

Using the package `utils`,


```{r}
sessionInfo()
```

Using devtools package

```{r}
session_info()
```

### A proposed workflow

The package `usethis` <https://usethis.r-lib.org/index.html> simplifies the setup of a project by:

* Creating a folder for the analysis
* Creating an RStudio project in the new folder
* Initializing a local git repo in that folder
* Opening a new RStudio session for the new project (so, working directory is in the new folder)

These workflows are well explained in [Happy Git and GitHub for the useR](https://happygitwithr.com/)

#### GitHub first (SSH)

1. Create a Github new repo.
1. Clone it as ssh. It searches for the key at: %HOME%/.ssh and must have name: id_rsa
1. Copy the ssh string in the clone with ssh
1. In RStudio, Create New project as git repo.
1. Copy the test in the git URL, chose the folder
1. You can commit any chnage to your local repo and push it later to remote.
1. Every time we push, it will ask for the passphrase.
1. To avoid it, we can run pageant and add the key we are using (It does not work)

#### GitHub first (HTTPS)

1. Create a Github new repo.
1. Clone it as https. 
1. Copy the https string in the clone with https
1. In RStudio, Create New project as git repo.
1. Copy the test in the git URL, chose the folder
1. You can commit any change to your local repo and push it later to remote.
1. The PAT (github token must passed to .Renviron, see how to check it below)

#### Using `usethis`

Three steps

*1* Create the new Rstudio project and folder. A new session of Rstusio will be open

```{r, eval = FALSE}
# 1. create folder and create a new Rproj there.
create_project(path = "C:/Users/josep.PORTATIL-ISABEL/Documents/R/...", open = TRUE, rstudio = TRUE)
```

*2* Create a local git repository. Rstudio will be rebooted to add the Git tab in Rstudio 

```{r, eval = FALSE}
# 2. Init a local git repo in that directory
use_git()
```

*3* Finally, create the GitHub repo. We can choose if connect via HTTPS (recommended) or ssh. Github repo will open. We can modifiy repo description

```{r, eval = FALSE}
#3. create a remote repo in github 
use_github()
```

We can checnk all the project setup with 

```{r}
usethis::git_sitrep()
```

